Flabby and injective objects in toposes

### 2022-07-06 paper-flabby-objects.tex
typo

\begin{defn}\begin{enumerate}
\item An abelian group~$I$ is \emph{divisible} if and only if for every
element~$x \in I$ and every natural number~$n \geq 1$, there is an element~$y \in I$
such that~$ny = x$.
\item An abelian group~$I$ \emph{satisfies the Baer criterion} if and only if,
for every ideal~$\aaa \subseteq \ZZ$, every ($\ZZ$-)linear map~$\aaa \to I$
admits an extension along the inclusion to a linear map~$\ZZ \to I$.
\end{enumerate}
\end{defn}

\begin{prop}
\label{prop:embed1}
\begin{enumerate}
\item Assuming the law of excluded middle, all divisible abelian groups satisfy
the Baer criterion.
\item Assuming Zorn's lemma, all abelian groups which satisfy the Baer criterion
are injective.
\end{enumerate}
\end{prop}

@@

Assuming Zorn's lemma, let~$I$ be an abelian group satisfying the Baer
criterion, let~$i : A \to B$ be a linear injection and let~$f : A \to I$ be a
linear map. The poset of partial extensions contains suprema of chains and
hence a maximal partial extension~$f_0 : A_0 \to I$ with~$A \subseteq A_0
\subseteq B$.

@@

\begin{proof}Let~$i : \aaa \hookrightarrow \ZZ$ be the inclusion of an ideal
and let~$f : \aaa \to A$ be a linear map as in the solid part of the following
commutative diagram:
\[ \xymatrix{
  \aaa \ar@{^{(}->}[r]^i\ar[d]\ar@/_2pc/[dd]_f & \ZZ \ar[d] \\
  \aaa^{++} \ar@{^{(}->}[r]^{i^{++}}\ar@{-->}[d] & \ZZ^{++}\ar@{-->}@/^/[ld]^{f'} \\
  (\QQ/\ZZ)^{++}
} \]
The given map~$f$ factors (uniquely) over the map~$\aaa \to \aaa^{++}$
since~$(\QQ/\ZZ)^{++}$ is a sheaf. The map~$i^{++}$ is still injective as
sheafification is exact. (The two top vertical maps need not be injective,
as~$\ZZ$ need not be separated for~$\nabla$.) Hence we are reduced to an
extension problem in the subtopos of~$\nabla$-sheaves, which is solvable
because this subtopos is Boolean (Proposition~\ref{prop:embed1}(1)). More
explicitly, the extension~$f'$ maps~$1 \in \ZZ$ to the gluing of
\begin{multline*}
  \{ 0 \,|\, \forall x \in \aaa. \nabla(x = 0) \} \cup {}\\
  \{ \tfrac{1}{d} v \,|\, d \geq 1, v \in [0,1),
    f(d) = \underline{v}, \forall x \in \aaa. \nabla(x \in \aaa)
    \Leftrightarrow \nabla(\exists n \in \ZZ. x = nd) \}.
\end{multline*}

@@

The second part is more involved and requires, similar to the proof of
Theorem~\ref{thm:enough-flabby-modules}, a strengthening of the Baer condition
where the required extensions are explicitly given. As a first step, let~$A$ be
an abelian group which is separated for a modal operator~$\nabla$ such
that~$\nabla(\varphi \vee (\varphi \Rightarrow \nabla\bot))$. The codomain of
the canonical map
\[ A \longrightarrow \prod_{g : A \to (\QQ/\ZZ)^{++}} (\QQ/\ZZ)^{++},\
  x \longmapsto (g(x))_g \]
is a product of abelian groups which functionally satisfy the Baer criterion by
Lemma~\ref{lemma:q-mod-z}(1) and hence satisfies the Baer condition itself.
Assuming Zorn's lemma, this map is injective by Lemma~\ref{lemma:q-mod-z}(2)
and~$\nabla$-separatedness of~$A$.

For the general case, let an abelian group~$A$ be given. Denoting
by~$(\cdot)^{+_x}$ the plus construction with respect to the modal
operator~$\nabla_x$ with~$\nabla_x\varphi \defeqv ((\varphi \Rightarrow x
= 0) \Rightarrow x = 0)$, the codomain of the canonical map
\[ A \longrightarrow \prod_{x \in A} \prod_{g : A \to (\QQ/\ZZ)^{+_x+_x}}
(\QQ/\ZZ)^{+_x+_x} \]
satisfies the Baer criterion and is injective as in the proof of
Theorem~\ref{thm:enough-flabby-modules}.
\end{proof}

>>>>

\begin{defn}\begin{enumerate}
\item An abelian group~$I$ is \emph{divisible} if and only if for every
element~$x \in I$ and every natural number~$n \geq 1$, there is an element~$y \in I$
such that~$ny = x$.
\item An abelian group~$I$ \emph{satisfies the Baer condition} if and only if,
for every ideal~$\aaa \subseteq \ZZ$, every ($\ZZ$-)linear map~$\aaa \to I$
admits an extension along the inclusion to a linear map~$\ZZ \to I$.
\end{enumerate}
\end{defn}

Trivially, every group satisfying the Baer condition is also divisible.

\begin{prop}
\label{prop:embed1}
\begin{enumerate}
\item Assuming the law of excluded middle, all divisible abelian groups satisfy
the Baer condition.
\item Assuming Zorn's lemma, all abelian groups which satisfy the Baer condition
are injective.
\end{enumerate}
\end{prop}

@@

Assuming Zorn's lemma, let~$I$ be an abelian group satisfying the Baer
condition, let~$i : A \to B$ be a linear injection and let~$f : A \to I$ be a
linear map. The poset of partial extensions of~$f$ contains suprema of chains and
hence a maximal partial extension~$f_0 : A_0 \to I$ with~$A \subseteq A_0
\subseteq B$.

@@

\begin{proof}Let~$i : \aaa \hookrightarrow \ZZ$ be the inclusion of an ideal
and let~$f : \aaa \to (\QQ/\ZZ)^{++}$ be a linear map as in the solid part of the following
commutative diagram:
\[ \xymatrix{
  \aaa \ar@{^{(}->}[r]^i\ar[d]\ar@/_2pc/[dd]_f & \ZZ \ar[d] \\
  \aaa^{++} \ar@{^{(}->}[r]^{i^{++}}\ar@{-->}[d] & \ZZ^{++}\ar@{-->}@/^/[ld]^{f'} \\
  (\QQ/\ZZ)^{++}
} \]
The given map~$f$ factors (uniquely) over the map~$\aaa \to \aaa^{++}$
since~$(\QQ/\ZZ)^{++}$ is a sheaf. The map~$i^{++}$ is still injective as
sheafification is exact. (The two top vertical maps need not be injective,
as~$\ZZ$ need not be separated for~$\nabla$.) Hence we are reduced to an
extension problem in the subtopos of~$\nabla$-sheaves, which is solvable
because this subtopos is Boolean (Proposition~\ref{prop:embed1}(1)). More
explicitly, the extension~$f'$ maps~$1 \in \ZZ$ to the gluing of
\begin{multline*}
  \{ 0 \,|\, \forall x \in \aaa. \nabla(x = 0) \} \cup {}\\
  \{ \tfrac{1}{d} v \,|\, d \geq 1, v \in [0,1),
    f(d) = \underline{v}, \forall x \in \aaa. \nabla(x \in \aaa)
    \Leftrightarrow \nabla(\exists n \in \ZZ. x = nd) \}.
\end{multline*}

@@

The second part is more involved and requires, similar to the proof of
Theorem~\ref{thm:enough-flabby-modules}, a strengthening of the Baer condition
where the required extensions are explicitly given. As a first step, let~$A$ be
an abelian group which is separated for a modal operator~$\nabla$ such
that~$\nabla(\varphi \vee (\varphi \Rightarrow \nabla\bot))$. The codomain of
the canonical map
\[ A \longrightarrow \prod_{g : A \to (\QQ/\ZZ)^{++}} (\QQ/\ZZ)^{++},\
  x \longmapsto (g(x))_g \]
is a product of abelian groups which functionally satisfy the Baer condition by
Lemma~\ref{lemma:q-mod-z}(1) and hence satisfies the Baer condition itself.
Assuming Zorn's lemma, this map is injective by Lemma~\ref{lemma:q-mod-z}(2)
and~$\nabla$-separatedness of~$A$.

For the general case, let an abelian group~$A$ be given. Denoting
by~$(\cdot)^{+_x}$ the plus construction with respect to the modal
operator~$\nabla_x$ with~$\nabla_x\varphi \defeqv ((\varphi \Rightarrow x
= 0) \Rightarrow x = 0)$, the codomain of the canonical map
\[ A \longrightarrow \prod_{x \in A} \prod_{g : A \to (\QQ/\ZZ)^{+_x+_x}}
(\QQ/\ZZ)^{+_x+_x} \]
satisfies the Baer condition and is injective as in the proof of
Theorem~\ref{thm:enough-flabby-modules}.
\end{proof}
